An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-26 EconGrowth.Rnw
Further edits on the data section. Also adding several additional papers to the bibliography file.

\usepackage{soul} %Provides strikethrough text
\usepackage{float} %Allows precise positioning of tables and figures within the text
\usepackage{amsmath} %Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow} % To create multirow tables.

@@

%++++++++++++++++++++++++++++++
\subsection{Quantification of Energy} \label{sec:Quantification_Of_Energy}
%++++++++++++++++++++++++++++++

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Sources} \label{sec:Data_Sources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To assess energy-augmented models of economic growth along the coordinates of analysis outlined in Section \ref{sec:Coordinates}, historical data for economic output ($Y$), capital stock ($K$), labor ($L$), thermal energy ($Q$), exergy ($X$), and useful work ($U$) is required. Subsections below describe sources for and calculations to obtain historical data. A final subsection presents the data in graphical form.

************* Matt ended here. ****************

@@

We take real gross domestic product to be an appropriate, if imperfect, proxy for output ($Y$). Gross domestic product data in real 2005 U.S. dollars for the United States, the United Kingdom, and Japan for 1980-2011 is from the World Bank's Global Economic Prospects (GEP) database (WORLD BANK, 2012). GDP data in real 2005 U.S. dollars for South Africa, China, Iran, Saudi Arabia, Tanzania, and Zambia for 1991-2011 is from the World Bank's Global Economic Prospects database (WORLD BANK, 2012).

%++++++++++++++++++++++++++++++
\subsection{Capital stock ($K$)} \label{sec:Capital_Stock_Data}
%++++++++++++++++++++++++++++++

Gross capital stock for the United States, the United Kingdom, and Japan for 1980-2011 in real 2005 U.S. dollars is from the World Bank GEP database (WORLD BANK, 2012). Gross capital stock data in real 2005 U.S. dollars for South Africa, China, Iran, Saudi Arabia, Tanzania, and Zambia for 1991-2011 is from the World Bank GEP database (WORLD BANK, 2012).

%++++++++++++++++++++++++++++++
\subsection{Labor and employment ($L$)} \label{sec:Labor_Data}
%++++++++++++++++++++++++++++++

Total employment and average hours worked per year per worker for the United States, the United Kingdom, and Japan for 1980-2011 is from the United States Bureau of Labor Statistics (BLS, 2011). Total hours worked is calculated by multiplying total employment by average hours worked per worker. The World Bank's WDI database (WORLD BANK, 2012) provides an employment to population ratio of people fifteen and older, total population, and a percentage of people under fifteen for South Africa, China, Iran, Saudi Arabia, Tanzania, and Zambia for 1991-2011. The population of people fifteen and older was calculated by multiplying the percentage of people under fifteen by total population, and then subtracting this from the total population. Total employment could be calculated by multiplying the employment to population ratio of people fifteen and older by the population of people fifteen and older. The labor input for South Africa, China, Iran, Saudi Arabia, Tanzania, and Zambia is based on total employment, not on total hours worked per year.

%++++++++++++++++++++++++++++++
\subsection{Thermal energy, exergy, and useful work ($E$)} \label{sec:Energy_Data}
%++++++++++++++++++++++++++++++

@@

%------------------------------
\subsubsection{Thermal energy ($Q$)} \label{sec:Q_Data}
%------------------------------

@@

%------------------------------
\subsubsection{Useful work ($U$)} \label{sec:U_Data}
%------------------------------

@@

Funding for Caleb Reese and Lucas Timmer was generously supplied by the Jack and Lois Kuipers Applied Mathematics Endowment and the Calvin College Alumni Association. Support for Matthew Kuperus Heun was provided by a Calvin College Research Followship. Funding for Martin de Wit was provided by *******. The authors thank Loren L. Heun (Western Michigan University) for insightful critique of the statistical analyses presented herein.
%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Tables}
\setcounter{table}{0} % Restarts table number counter at 0 in this section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\usepackage{soul} % Provides strikethrough text
\usepackage{float} % Allows precise positioning of tables and figures within the text
\usepackage{amsmath} % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow} % To create multirow tables
\usepackage{hyperref} % To create hyperlinks in the paper

@@

%++++++++++++++++++++++++++++++
\subsection{Quantification of Energy} 
\label{sec:Quantification_of_Energy}
%++++++++++++++++++++++++++++++

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Historical Data}
\label{sec:Historical_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To assess energy-augmented models of economic growth along the coordinates of analysis outlined in Section \ref{sec:Coordinates}, historical data for economic output ($Y$), capital stock ($K$), labor ($L$), thermal energy ($Q$), exergy ($X$), and useful work ($U$) are required. Subsections below describe sources for and calculations to obtain historical data. A final subsection presents the historical data in graphical form.

@@

We measure economic output ($Y$) by real gross domestic product (GDP). GDP data in real 2005 U.S. dollars for developed countries (1980-2011) and all other countries (1991-2011) was obtained from the World Bank's Global Economic Prospects (GEP) database \citep{WorldBankGEP:2013a}.

%++++++++++++++++++++++++++++++
\subsection{Capital Stock ($K$)} 
\label{sec:Capital_Stock_Data}
%++++++++++++++++++++++++++++++

Gross capital stock in real 2005 U.S. dollars for developed countries (1980-2011) and all other countries (1991-2011) was obtained from the World Bank's Global Economic Prospects (GEP) database \citep{WorldBankGEP:2013a}.

%++++++++++++++++++++++++++++++
\subsection{Labor and Employment ($L$)} 
\label{sec:Labor_Data}
%++++++++++++++++++++++++++++++

Total employment and average hours worked per year per worker for the United States, the United Kingdom, and Japan for 1980-2011 was obtained from the United States \citet{BLS:2011a}. Total hours worked was calculated by multiplying total employment by average hours worked per worker. ************* Matt ended here. **************** The World Bank's WDI database (WORLD BANK, 2012) provides an employment to population ratio of people fifteen and older, total population, and a percentage of people under fifteen for South Africa, China, Iran, Saudi Arabia, Tanzania, and Zambia for 1991-2011. The population of people fifteen and older was calculated by multiplying the percentage of people under fifteen by total population, and then subtracting this from the total population. Total employment could be calculated by multiplying the employment to population ratio of people fifteen and older by the population of people fifteen and older. The labor input for South Africa, China, Iran, Saudi Arabia, Tanzania, and Zambia is based on total employment, not on total hours worked per year.

%++++++++++++++++++++++++++++++
\subsection{Thermal Energy, Exergy, and Useful Work ($E$)} 
\label{sec:Energy_Data}
%++++++++++++++++++++++++++++++

@@

%------------------------------
\subsubsection{Thermal Energy ($Q$)} 
\label{sec:Q_Data}
%------------------------------

@@

%------------------------------
\subsubsection{Useful Work ($U$)} 
\label{sec:U_Data}
%------------------------------

@@

Funding for Caleb Reese and Lucas Timmer was generously supplied by the Jack and Lois Kuipers Applied Mathematics Endowment and the Calvin College Alumni Association. Support for Matthew Kuperus Heun was provided by a Calvin College Research Followship. Funding for Martin de Wit was provided by *******. The authors thank Loren L. Heun (Western Michigan University) for insightful critique of the statistical analyses presented herein.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reproducible Research}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the spirit of Reproducible Research \citep{Gandrud:2013vx}, all data, spreadsheets, \texttt{R} code \citep{R}, and other materials associated with this paper can be found at \url{https://github.com/MatthewHeun/Econ-Growth-R-Analysis}.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Covariance Tables}
\label{sec:Covariance_Tables}
\setcounter{table}{0} % Restarts table number counter at 0 in this section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
